---
title: "Estimate Daily Customer Level Usage and Usage Shifting"
author: "Yu Zhou"
date: "May 27, 2016"
output: html_document
---


```{r, include=FALSE}
library(ggplot2)
library(reshape2)
library(plyr)
library(sqldf)
library(forecast)
library(zoo)
rw_with_et <- read.csv("~/Desktop/DataKind/Data/rw_with_et.csv")
ProductionConsumption <- read.csv("~/Desktop/DataKind/Data/ProductionConsumption.csv")
```

```{r, echo=FALSE, include=FALSE}
#this is for ARIMA model
ProductionConsumption$Date<-as.Date(ProductionConsumption$Date,"%m/%d/%y")
acf(ProductionConsumption$AmountDelivered,type="partial")
auto.arima(ProductionConsumption$AmountDelivered, stepwise=FALSE, approximation=FALSE,seasonal=TRUE)
```


From the production data set, I use the aggregated delivery amount in the past years to make a strong assumption about individual user's day to day water usage amount pattern. I assume that each individual follows the aggregated delivery amount pattern. I use local averages of 15 days to generate a linear estimate of individual user's day to day water usage amount pattern. To avoid much data manipulation due to date matching, I only focus on data from 2013. 


```{r}
temp<-rollmean(ProductionConsumption$AmountDelivered, 15)
temp<-c(ProductionConsumption$AmountDelivered[1:7],temp,ProductionConsumption$AmountDelivered[4226:4232])
ProductionConsumption$LocalAverage<-temp
#ggplot()+ geom_line(data=ProductionConsumption, aes(x=Date, y=AmountDelivered, color="AmountDelivered"))+ #geom_line(data=ProductionConsumption, aes(x=Date, y=LocalAverage,color="LocalAverage")) 
pc_2013<-sqldf("select * from ProductionConsumption where Year=2013 ")
ggplot()+ geom_line(data=pc_2013, aes(x=Date, y=AmountDelivered, color="AmountDelivered"))+ geom_line(data=pc_2013, aes(x=Date, y=LocalAverage,color="LocalAverage")) +ggtitle("The 2013 Aggregated Water Usage Pattern")
```


From previous analysis, user id 64081 is identifie as the top offender. I want to construct a daily water usage for 64081. I know 64081's total montly water usage in 2013, now I construct 64081's daily water usage in 2013. 


```{r}
et_64081<-sqldf("select usage_ccf, usage_month as Month, usage_year from rw_with_et where cust_id_from_utility=64081 and usage_year=2013")
pc_2013_LocalAverage<-sqldf("select LocalAverage,Month, Date from pc_2013")
pc_2013_LocalAverage$index<-c(1:length(pc_2013_LocalAverage$LocalAverage))
daily_use_64081<-sqldf("select * from pc_2013_LocalAverage left join et_64081 on et_64081.Month=pc_2013_LocalAverage.Month")
daily_use_64081<-daily_use_64081[,-2]
daily_use_64081<-sqldf("select * from daily_use_64081 order by usage_year, Month")
MonthSum<-sqldf("select Month, sum(LocalAverage) as Sum from daily_use_64081 group by Month")
daily_use_64081<-sqldf("select * from daily_use_64081 left join MonthSum on daily_use_64081.Month=MonthSum.Month")
daily_use_64081$adjust<-(daily_use_64081$usage_ccf/daily_use_64081$Sum)
daily_use_64081$LocalAverage<-daily_use_64081$LocalAverage*daily_use_64081$adjust
daily_use_64081$ImagineOne<-daily_use_64081$LocalAverage*1.5
daily_use_64081$ImagineTwo<-daily_use_64081$LocalAverage*2

ggplot()+ geom_line(data=daily_use_64081, aes(x=Date, y=LocalAverage,color="Estimated Usage"))+ggtitle("64081's Estimated Daily Water Usage in 2013")
```

Because I assume all users follow the same pattern, therefore I can create estimated water usage patterns for other users. 

```{r}
ggplot()+ geom_line(data=daily_use_64081, aes(x=Date, y=LocalAverage,color="64081")) +geom_line(data=daily_use_64081, aes(x=Date, y=ImagineOne,color="A"))+geom_line(data=daily_use_64081, aes(x=Date, y=ImagineTwo,color="B"))+ggtitle("Estimated Daily Water Usage in 2013")
```

There are obvious cliffs at the begining/end of each month. Real individual usage lines should not have cliffs at the begining/end of each month. The next step is to experiment with usage shifting. The usage shifting is constrained within each month.


**Usage Shifting**

We combine 64081, ImagineOne and ImagineTwo into Total. Total is the toal water usage we want to control. Because of the seasonality of water usage, I think we can identify peak months and then find opportunities to manage demand. Here, with just eyeballing, I want to identify June, Augest and Octorber as peak months. 

```{r}
daily_use_64081$Total<-daily_use_64081$LocalAverage+daily_use_64081$ImagineOne+daily_use_64081$ImagineTwo
ggplot()+ geom_line(data=daily_use_64081, aes(x=Date, y=LocalAverage,color="64081")) +geom_line(data=daily_use_64081, aes(x=Date, y=ImagineOne,color="A"))+geom_line(data=daily_use_64081, aes(x=Date, y=ImagineTwo,color="B"))+geom_line(data=daily_use_64081, aes(x=Date, y=Total,color="Total"))+ggtitle("Estimated Daily Water Usage in 2013")
```

Because our usage pattern are estimated, it is not valid to manage demand across months. In the next step, I will only focus on June. 

```{r}
daily_use_64081<-daily_use_64081[,-5]
june_use<-sqldf("select * from daily_use_64081 where Month=6")
ggplot()+ geom_line(data=june_use, aes(x=Date, y=LocalAverage,color="64081")) +geom_line(data=june_use, aes(x=Date, y=ImagineOne,color="A"))+geom_line(data=june_use, aes(x=Date, y=ImagineTwo,color="B"))+geom_line(data=june_use, aes(x=Date, y=Total,color="Total"))+ggtitle("Estimated Daily Water Usage in 2013 June")
```


An intuitive approach, find the highest peak, if a 3 day switch benefits, then switch, otherwise, exam the second highest peak. After a switch, calculate the total again. 

```{r}
june_use<-june_use[order(june_use$Date,decreasing =FALSE),]
june_use$diff<-c(diff(june_use$Total,3),c(NA,NA,NA))
for (i in 1:10){
  if (june_use$diff[june_use$Total==sort(june_use$Total,TRUE)[i]]<0){
    Date<-june_use$Date[june_use$Total==sort(june_use$Total,TRUE)[i]]
    temp<-june_use$LocalAverage[june_use$Date==Date]
    NewDate<-Date+3
    june_use$LocalAverage[june_use$Date==Date]<-june_use$LocalAverage[june_use$Date==NewDate]
    june_use$LocalAverage[june_use$Date==NewDate]<-temp
    june_use$Total<-june_use$LocalAverage+june_use$ImagineOne+june_use$ImagineTwo
    june_use$diff<-c(diff(june_use$Total,3),c(NA,NA,NA))
  }    
  print(Date)
}

ggplot()+ geom_line(data=june_use, aes(x=Date, y=LocalAverage,color="64081")) +geom_line(data=june_use, aes(x=Date, y=ImagineOne,color="A"))+geom_line(data=june_use, aes(x=Date, y=ImagineTwo,color="B"))+geom_line(data=june_use, aes(x=Date, y=Total,color="Total"))+ggtitle("Adjusted Daily Water Usage in 2013 June")

```

The peak is lower a little bit. 